\section{Conclusion and Future Work}
\label{sec:conclusion}

This paper presents Charm, a novel system that improves battery life and range
in LP-WANs clients. Charm achieves this through a mechanism that pools
together weak received signals across multiple gateways at the cloud in order
to jointly decode them. In doing so, Charm overcomes multiple challenges,
including a hardware-software design to detect weak signals at the gateway,
and ensure scalability at the cloud. A pilot evaluation of Charm on a network
of twelve LoRaWAN gateways serving a large neighborhood of a major U.S. city
demonstrates a large improvement in coverage and client battery-life.

An interesting side-benefit of Charm is its impact on scalability of the
network overall. Given that Charm improves coverage, one might expect a large
number of collisions from transmitters who newly gain coverage with existing
ones. Counter-intuitively, this is not the case because Charm allows devices
across the board to transmit at faster data rates, increasing available air
time in the network. Our future work will explore further improving network
scalability along two dimensions: (1) First, a full-scale distributed MIMO
system atop LP-WAN in the cloud, that can also handle collisions from a large
number of clients. (2) Second, offloading of TV whitespace spectrum at peak
demand, based on an FCC license recently granted to our university.



% This paper presents a novel location-aware network management system for
% LoRa-class LP-WANs operating in unlicensed and whitespace frequencies. It
% presents an RF-based localization system that stitches together information
% across frequencies to improve positioning accuracy of LoRa clients, even
% without access to their channel state information. We build on the
% localization framework to build a resource allocation system for LP-WAN that
% efficiently allocates wireless resources subject to FCC's regulatory
% constraints.  Our network-management system  is designed to be
% location-aware, exploiting live measurements to identify and respond to
% interference and help network managers plan changes to deployment. Our
% system was implemented and deployed in a large university campus and results
% from proof-of-concept experiments and large-scale trace-driven simulations
% are presented. In the future, we aim to expand our deployment to incorporate
% transmissions on the white spaces and are currently engaged in efforts to
% procure the relevant licenses. We also intend to integrate smart sensor
% devices currently deployed on-campus with LoRaWAN radios to evaluate our
% network management platform live and at-scale.