% \documentclass[sigconf, anonymous]{acmart}
\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{graphicx}
% Note: do not use url & hyperref packages.
% They do not pass the IEEE pdfxpress checks
% \usepackage{url}
% \usepackage{hyperref}
% \usepackage{subfigure}
\usepackage{balance}
\usepackage{amssymb,amsmath}
% \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{algorithm2e}
\usepackage{algpseudocode}

\usepackage{todonotes} % get rid of this later
\usepackage{subfig}
\usepackage{wrapfig}
\newcommand{\compactimg}{\vspace{-14pt}}
\newcommand{\tableref}[1]{Table~\ref{tab:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\algoref}[1]{Algorithm~\ref{algo:#1}}

\newcommand{\bob}[1]{\textcolor{red}{{\bf from Bob: #1}}}
\newcommand{\cmt}[2]{{\color{blue} #1} {\color{red} [#2]}}

\newenvironment{tightitem}{\begin{itemize}\setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}{\end{itemize}}

\begin{document}

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
% \setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
% \acmDOI{None}

% ISBN
% \acmISBN{}

%Conference
% \acmConference[IPSN]{IPSN}{2018}{Porto} 
% \acmYear{2018}
% \copyrightyear{2018}

% \acmPrice{xx.00}

\title{ Charm: Exploiting Geographical Diversity Through Coherent Combining in Low-Power Wide-Area Networks }

%\titlenote{Produces the permission block, and
%  copyright information}
% \subtitle{Extended Abstract}
% \subtitlenote{The full version of the author's guide is available as
%   \texttt{acmart.pdf} document}

% \author{Paper \# 122}

% \author{
% 	\IEEEauthorblockN{
% 		Adwait Dongare,
% 		Revathy Narayanan,
% 		Akshay Gadre,
% 		Anh Luong,
% 		Artur Balanuta,\\
% 		Swarun Kumar,
% 		Anthony Rowe\\}
% 	\IEEEauthorblockA{
% 		Electrical and Computer Engineering Department,\\
% 		Carnegie Mellon University, Pittsburgh PA, USA\\
% 	}
% }

\author{Adwait Dongare, Revathy Narayanan, Akshay Gadre, Anh Luong, Artur Balanuta, \\ Bob Iannucci, Swarun Kumar, Anthony Rowe }
\affiliation{%
 \institution{Carnegie Mellon University}
 \department{Electrical and Computer Engineering}
 % \streetaddress{5000 Forbes Ave.}
 \city{Pittsburgh} 
 \state{Pennsylvania} 
 % \postcode{12345}
}
\email{ [adongare, revathyn, agadre, aluong2, arturb, roberti, swarunk, agr] @andrew.cmu.edu }

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Dongare et al.}


\begin{abstract}


Low-Power Wide Area Networks (LP-WANs) are an emerging wireless platform which
can support battery-powered devices lasting 10-years, while communicating at
low data-rates to gateways several miles away. Yet, despite the expected
high-density of LP-WAN gateways in future cities, not all devices will
experience the promised 10-year battery life, particularly in urban spaces. A
large number of devices that are deep inside buildings, or in remote
neighborhoods would suffer severe battery-drain due to extended transmissions
at the slowest data rate to reach even the closest gateway.

This paper presents Charm, a system that enhances both battery life and
coverage of LP-WAN clients in large urban deployments. Charm allows multiple
LoRaWAN gateways to pool their received signals in the cloud, coherently
combining them to detect even the weakest signal that is not decodable at any
individual gateway. Charm achieves this through a novel hardware and software
design at the gateway that carefully detects which chunks of the received
signal need to be sent to the cloud, thereby saving uplink bandwidth. We
demonstrate how our solution scales to decoding weak transmissions at
city-scale, by identifying the set of gateways whose signals need to be
coherently combined over time. In evaluations over a test network and through
trace-driven simulations from a large LoRaWAN deployment in Pittsburgh, Charm
demonstrates a significant gain of 3$\times$ in range and 4$\times$ in client
battery-life.

\end{abstract}

%used to add page numbers
\settopmatter{printfolios=true}

\maketitle

\input{tex/1-introduction.tex}
\input{tex/2-related.tex}
\input{tex/3-background.tex}
\input{tex/4-arch.tex}
\input{tex/5-gateway.tex}
\input{tex/6-cloud.tex}
\input{tex/7-implementation.tex}
\input{tex/8-evaluation.tex}
\input{tex/9-conclusion.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{references} 

\end{document}
